---
title: "Untitled"
author: "Tourism Economics"
date: "April 15, 2016"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}

library(arlodr)
library(zoo, warn.conflicts=FALSE)
library(xts, warn.conflicts=FALSE)
library(dplyr, warn.conflicts=FALSE)
library(tidyr, warn.conflicts=FALSE)
library(ggplot2, warn.conflicts=FALSE)
library(knitr)
library(readr)


```

```{r}

# load certain meta data
ffile <- c("~/Project/R projects/bts_2016/input_data/T100_D_MARKET_US_CARRIER_ONLY_META")
l_airport_id <- read_csv(paste0(ffile, "/", "L_AIRPORT_ID.csv"), col_names=TRUE)
colnames(l_airport_id) <- tolower(colnames(l_airport_id))

ffile <- c("~/Project/R projects/bts_2016/input_data/T100_D_MARKET_US_CARRIER_ONLY_META")
l_city_market_id <- read_csv(paste0(ffile, "/", "L_CITY_MARKET_ID.csv"), col_names=TRUE)
colnames(l_city_market_id) <- tolower(colnames(l_city_market_id))

# load manual crosswalk
# the dest_city_market_id column is the part that is stable. The names can change. So use the
# codes for lookups.
ffile <- c("~/Project/R projects/bts_2016/input_data/")
manual_cw_from_bts_cityid <- read_csv(paste0(ffile, "/", "manual_cw_from_bts_cityid.csv"), col_names=TRUE)

# load census metro list from fred
ffile <- c("~/Project/R projects/bts_2016/input_data/")
m_cen <- read_csv(paste0(ffile, "/", "m_cen.csv"), col_names=TRUE)

# load data
load("~/Project/R projects/bts_2016/output_data/t100d_uscar.Rdata")

```


```{r echo=FALSE}

data1 <- t100d_uscar

# add certain meta data
label_airport_id <- l_airport_id %>%
  rename(dest_airport_name = description)

#############

# add certain meta data
label_city_market_id <- l_city_market_id %>%
  rename(dest_city_name = description)

# drop the dest_city_name because it may be old, and
# instead we can use the one from the label_city_market_id file
manual_cw_from_bts_cityid <- manual_cw_from_bts_cityid %>%
  select(-dest_city_name)

data_work <- data1 %>%
  left_join(label_airport_id, by = c("dest_airport_id" = "code")) %>%
  # remove the existing dest_city_name because I noticed that Dallas seemed to come up with two different
  # text names, under one code, so I thought I'd just merge on a recent
  # list of destination city names
  select(-dest_city_name) %>%
  left_join(label_city_market_id, by = c("dest_city_market_id" = "code")) %>%
  left_join(manual_cw_from_bts_cityid, by = c("dest_city_market_id")) %>%
  # rename the passenger column
  rename(pax = passengers)

```


Generate a list of cities, which was useful for creating a manual crosswalk.
```{r}

city_by_volume <- data_work %>%
  filter(year == 2015) %>%
  group_by(dest_city_market_id, dest_city_name) %>%
  summarise(pax = sum(pax)) %>%
  ungroup() %>%
  arrange(desc(pax))
kable(head(city_by_volume,50))

write.csv(city_by_volume, "~/Project/R projects/bts_2016/output_data/city_by_volume.csv", row.names=FALSE)
```



# Experimenting a bit with Atlanta
```{r}

temp1 <- data_work %>%
  filter(unique_carrier == "DL") %>%
  group_by(dest_airport_name, year, month) %>%
  summarise(pax = sum(pax))

atlanta <- temp1 %>%
  filter(year == 2014) %>%
  filter(dest_airport_name == "Atlanta, GA: Hartsfield-Jackson Atlanta International")

kable(atlanta)
```

For December 2014, I keep getting 2,457,492. Same as Excel, and different from the web page which shows
2,456,412.

```{r}

atlanta <- data_work %>%
  filter(dest_airport_id == 10397) %>%
  filter(unique_carrier == "DL") %>%
  group_by(year, month) %>%
  summarise(pax = sum(pax))

kable(tail(atlanta, 18))

```

```{r}

atlanta <- data_work %>%
  filter(dest_city_market_id == 30397) %>%
  filter(unique_carrier == "DL") %>%
  group_by(year, month) %>%
  summarise(pax = sum(pax))

kable(tail(atlanta, 18))

```


```{r}
data_1 <- data_work %>%
  group_by(dest_city_market_id, dest_city_name, year, month) %>%
  summarise(pax = sum(pax))

atlanta <- data_1 %>%
  filter(year == 2015) %>%
  filter(dest_city_name == "Atlanta, GA (Metropolitan Area)")

kable(atlanta)
```

Use the manual crosswalk. For those areas that have an area_name_simp, see which airports fall into that area, while also seeing how the dest_city_names relate.
```{r}

temp3 <- data_work %>%
  filter(! is.na(area_name_simp)) 

air_by_area_simp  <- distinct(select(temp3, dest_city_name, area_name_simp, dest_airport_name)) %>%
  arrange(dest_city_name, area_name_simp, dest_airport_name)

kable(air_by_area_simp)


##############
#
# modify by making a few changes

# pull San Jose airport out of San Francisco to a San Jose, CA area_name_simp
data_work_mod1 <- data_work %>%
  mutate(area_name_simp = ifelse(dest_airport_id == 14831, "San Jose, CA", area_name_simp)) 

# Also, created a separate Orange County
temp_orgca <- data_work_mod1 %>%
  filter(dest_airport_id == 14908) %>%
  mutate(area_name_simp = "Orange County, CA")

data_work_mod2 <- rbind(data_work_mod1, temp_orgca)
  




```

## Take another look after the changes
As a result of the changes, the San Jose airport and the John Wayne airport in Orange County now
both have corresponding area_name_simp entries. The summary table still shows that from a bls destin_city_name
perspective, both are still in the same destin_city_name, but that isn't being used to aggregate the data
```{r}
## take another look after the changes

temp3 <- data_work_mod2 %>%
  filter(! is.na(area_name_simp)) 

air_by_area_simp  <- distinct(select(temp3, dest_city_name, area_name_simp, dest_airport_name)) %>%
  arrange(dest_city_name, area_name_simp, dest_airport_name)

kable(air_by_area_simp)
```


# Create output table
```{r}

# calculate a total US
data_tot <- data_work_mod2 %>%
  # this would be double counting, so first excluded Orange County
  # which is still also in LA
  filter(area_name_simp != "Orange County, CA") %>%
  group_by(year, month) %>%
  summarise(pax = sum(pax)) %>%
  ungroup() %>%
  mutate(area_sh = "usxxx") %>%
  mutate(area_name_simp = "United States") %>%
  mutate(date = paste(year, month, "01", sep="-")) %>%
  mutate(date = as.Date(date)) %>%
  select(-year, -month)

data_4 <- data_work_mod2 %>%
  filter(!is.na(area_name_simp)) %>%
  group_by(area_name_simp, year, month) %>%
  summarise(pax = sum(pax)) %>%
  ungroup() %>%
  ungroup()

data_5 <- data_4 %>%
  left_join(m_cen, by = "area_name_simp") %>%
  # fix Orange County which isn't in m_cen
  mutate(area_sh = ifelse(area_name_simp == "Orange County, CA", "orgca", area_sh)) %>%
  select(area_name_simp, area_sh, year, month, pax) %>%
  mutate(date = paste(year, month, "01", sep="-")) %>%
  mutate(date = as.Date(date)) %>%
  select(-year, -month) %>%
  select(date, everything()) 

# add on total
hold_t100d_uscar <- rbind(data_5, data_tot) %>%
  # add columns that describes the data
  # domestic flights
  mutate(flight_type = "d") %>%
  # us carriers
  mutate(carriers = "uscar") 

  
look <- hold_t100d_uscar %>%
  filter(area_sh == "usxxx")
ggplot(data=look, aes(x=date, y=pax)) + 
  geom_line()

look <- hold_t100d_uscar %>%
  filter(area_sh == "phlpa")
ggplot(data=look, aes(x=date, y=pax)) + 
  geom_line()

look <- hold_t100d_uscar %>%
  filter(area_sh == "wshdc")
ggplot(data=look, aes(x=date, y=pax)) + 
  geom_line()

look <- hold_t100d_uscar %>%
  filter(area_sh == "sttwa")
ggplot(data=look, aes(x=date, y=pax)) + 
  geom_line()

look <- hold_t100d_uscar %>%
  filter(area_sh == "snfca")
ggplot(data=look, aes(x=date, y=pax)) + 
  geom_line()

look <- hold_t100d_uscar %>%
  filter(area_sh == "dlltx")
ggplot(data=look, aes(x=date, y=pax)) + 
  geom_line()

look <- hold_t100d_uscar %>%
  filter(area_sh == "hsttx")
ggplot(data=look, aes(x=date, y=pax)) + 
  geom_line()

look <- hold_t100d_uscar %>%
  filter(area_sh == "snjca")
ggplot(data=look, aes(x=date, y=pax)) + 
  geom_line()

look <- hold_t100d_uscar %>%
  filter(area_sh == "orgca")
ggplot(data=look, aes(x=date, y=pax)) + 
  geom_line()

save(hold_t100d_uscar, file="~/Project/R projects/bts_2016/output_data/hold_t100d_uscar.Rdata")

```

